Converging High-Throughput and High-Performance Computing: A Case Study

### 2017-04-03 PandaEvolution.tex

% Titan’s backfill functionality offers the opportunity to avoid the overhead of
% queue wait times without using pilot abstraction. Backfill availability is the
% number of node-hours (or core-hours) that cannot be used for a certain amount
% of time by any of the jobs already queued on Titan: All queued jobs are either
% too large or their walltime is too long. At any point in time, Titan’s Moab
% scheduler can be queried for backfill availability. Based upon the result of
% this query, a job can be shaped to request no more than the backfill
% availability. As such, when submitted this job is usually scheduled
% immediately, spending almost no time in the queue.

Titan’s backfill functionality offers the opportunity to avoid the overhead of
queue wait times without using pilot abstraction~\cite{maui_backfill_url}.
Backfill availability is the number of node-hours (or core-hours) that cannot be
utilized by regular jobs queued on Titan at a specific point in time.  It is
important to note that backfill availability is computed retrospectively and is
determined by Titan's operation team. Backfill availability represents the
theoretical maximum of node-hours that could be utilized by ATLAS (or any other
project) that submit under backfill mode.

Compared to pilots, backfill has the disadvantage of limiting the amount of work
nodes that can be requested. Pilots are normal jobs: they can request as many
worker nodes for as much time as a queue can offer. On the contrary, jobs sized
on the basis of backfill information availability depend on the number of worker
nodes that cannot be given to any other job in the Titan's queue at that moment
in time.

@@

Backfill availability is typically a small fraction of the total capacity  of a
resource. Notwithstanding,  given the size of Titan this translates into a
substantial capacity. Every year, about 10\% of Titan's capacity remains
unused~\cite{barker2016us}, corresponding to an average of 30,000 unused cores.
This equals approximately 270M core hours per year, roughly 10\% of the overall
capacity of WLCG.

>>>>

% Titan’s backfill functionality offers the opportunity to avoid the overhead of
% queue wait times without using pilot abstraction. Backfill availability is the
% number of node-hours (or core-hours) that cannot be used for a certain amount
% of time by any of the jobs already queued on Titan: All queued jobs are either
% too large or their walltime is too long.

The backfill optimization of Titan's Moab scheduler offers the opportunity to
avoid the overhead of queue wait times without using pilot
abstraction~\cite{maui_backfill_url}. This optimization enables Moab to start
low-priority jobs as long as they do not delay higher priority jobs,
independently on whether the low-priority jobs were queued after the
high-priority jobs.

When the backfill optimization is enabled, users can interrogate Moab about what
number of worker nodes and walltime would be available to a low-priority job at
that moment in time. If a job is immediately submitted to Titan with that number
of worker nodes and walltime, chances are that Moab will immediately scheduled
it, reducing its queue time to a minimum. In this paper, we call this number of
worker nodes and walltime `backfill availability'.

% It is important to note that backfill availability is computed retrospectively
% and is determined by Titan's operation team. Backfill availability represents
% the theoretical maximum of node-hours that could be utilized by ATLAS (or any
% other project) that submit under backfill mode.

Compared to pilots, backfill has the disadvantage of limiting the amount of
worker nodes that can be requested. Pilots are normal jobs: they can request as
many worker nodes for as much time as a queue can offer. On the contrary, jobs
sized on the basis of backfill availability depend on the number of worker nodes
that cannot be given to any other job in the Titan's queue at that moment in
time.

@@

At any point in time, backfill availability is typically a small fraction of the
total capacity of a resource. Notwithstanding, given the size of Titan this
translates into a substantial capacity. Every year, about 10\% of Titan's
capacity remains unused~\cite{barker2016us}, corresponding to an average of
30,000 unused cores (excluding GPU cores). This equals approximately 270M core
hours per year, roughly 10\% of the overall capacity of WLCG.
